Multiple Dice Working as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs

### 2015-04-14 doc/journal/fpgapaperfinal/FPGAPaper/fpga2014paper.tex
Two minor updates by Vaughn (Virtex Ultrascale is shipping and our partition assignment is suboptimal only for >2 dice) plus added cover letter.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4595 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

The improvement in the capacity of 2.5D FPGAs over conventional ones is very significant. Xilinx's largest 28 nm interposer-based FPGA contains 4 FPGA dice and 1.954 million logic cells~\cite{xilinx7series}, while the largest non-interposer Virtex-7 die has 979k logic cells and Altera's largest 28 nm FPGA has 952k logic elements~\cite{stratixV}. Even though all these FPGAs use a 28 nm process, silicon interposer technology allows the creation of FPGAs with twice the resources possible on even an extremely large single die. Xilinx's recently announced Virtex UltraScale line of 20 nm FPGAs makes even more aggressive use of interposers; the largest monolithic FPGA in this family contains 941k logic cells, while the four largest capacity FPGAs are all built with silicon interposers and span a range from 1.25 to 4.43 million logic cells~\cite{virtexultrascale}.

>>>>

The improvement in the capacity of 2.5D FPGAs over conventional ones is very significant. Xilinx's largest 28 nm interposer-based FPGA contains 4 FPGA dice and 1.954 million logic cells~\cite{xilinx7series}, while the largest non-interposer Virtex-7 die has 979k logic cells and Altera's largest 28 nm FPGA has 952k logic elements~\cite{stratixV}. Even though all these FPGAs use a 28 nm process, silicon interposer technology allows the creation of FPGAs with twice the resources possible on even an extremely large single die. Xilinx's latest (Virtex UltraScale, 20 nm) line of FPGAs makes even more aggressive use of interposers; the largest monolithic FPGA in this family contains 941k logic cells, while the four largest capacity FPGAs are all built with silicon interposers and span a range from 1.25 to 4.43 million logic cells~\cite{virtexultrascale}.

### 2015-04-14 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Two minor updates by Vaughn (Virtex Ultrascale is shipping and our partition assignment is suboptimal only for >2 dice) plus added cover letter.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4595 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

\subsection{Partitioner Stage}\label{sec:partitioner_stage}
As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow. The first possibility is to partition the netlist of primitive elements (LUTs, FFs, RAM slices, etc.) before packing into the larger function blocks such as logic blocks and (32 kb) RAM blocks. The partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block. We updated the VPR clustering algorithm to respect this new constraint. The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist. Each partition is then assigned (randomly) to a separate FPGA die. Our use of random assignment is not optimal, and an interesting area for future work is to explore keeping strongly connected partitions in adjacent dice. The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines. These two partitioning flow options can be combined -- if partitioning is applied before packing, one can also pass the partition information to the placement stage. The post-packing placement constraints can be derived from the primitive block partitioning constraints, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block. There are, therefore, four possible partitioning flows:

>>>>

\subsection{Partitioner Stage}\label{sec:partitioner_stage}
As shown in Figure \ref{fig:partitioning_cad_flow}, we explore the effect of circuit partitioning at two different stages of the CAD flow. The first possibility is to partition the netlist of primitive elements (LUTs, FFs, RAM slices, etc.) before packing into the larger function blocks such as logic blocks and (32 kb) RAM blocks. The partitioning constraints must be respected by the packer; it must not pack two primitives assigned to different partitions (dice) into the same function block. We updated the VPR clustering algorithm to respect this new constraint. The second possible way to incorporate partitioning is after packing, where partitioning is applied to the function block (cluster-level) netlist. Each partition is then assigned (randomly) to a separate FPGA die. Our use of random assignment is sub-optimal for systems with more than two dice, and an interesting area for future work is to explore keeping strongly connected partitions in adjacent dice. The partitioning constraints must then be respected by the placer; it is free to move function blocks around within a die, but not across interposer cutlines. These two partitioning flow options can be combined -- if partitioning is applied before packing, one can also pass the partition information to the placement stage. The post-packing placement constraints can be derived from the primitive block partitioning constraints, with the valid placement region for a given function block being equal to the intersection of the valid placement regions of the primitives packed into that function (clustered) block. There are, therefore, four possible partitioning flows:
